---
title: "01-read_the_dataset"
output: html_notebook
---



```{r  rows.print=25}
# read the well data from a CSV file
well_data <- read.csv("oilfield_25w_dataset-01.csv", stringsAsFactors = FALSE)
well_data
```



```{r}
well_data
```


```{r}
# start by reading one well
well_name_01 <- well_data[1, "Wellname"]
well_name_01
```

```{r}
require(digest)

digest(well_name_01, algo = "crc32", serialize = FALSE)
```


```{r}
# read a second well
well_name_02 <- well_data[2, "Wellname"]   # get the well name from the 2nd row
well_name_02   # show the well
digest(well_name_02, algo = "crc32", serialize = FALSE)  # anonymize the well name
```

```{r}
# let's  do the last well in the table: No. 25
well_name_25 <- well_data[25, "Wellname"]   # get the well name from the 2nd row
well_name_25   # show the well
digest(well_name_25, algo = "crc32", serialize = FALSE)  # anonymize the well name
```

We can continue doing this all day long but becomes tedious.

```{r}
require(digest)
# creating a function to anonymize 
# param x     the column/variable to anonymize
# param algo  the hashing algorithm. options: crc32, sha1, sha256, sha512, etc.
anonymize <- function(x, algo = "crc32") {
    hashes <- vapply(unique(x), function(object)
                digest(object, algo = algo, serialize = FALSE), 
                FUN.VALUE = "", USE.NAMES = TRUE)
    
    unname(hashes[x])
}
```

**Notes: **
* serializing requires that the variables or columns are characters not factors, thay why we used the option `stringsAsFactors = FALSE` when we imported the data above.
* In this example we are using the hashing algorithm `crc32` but we could use another one stronger such as `sha1`.
* Be sure that you are using the option `serialize = FALSE` in digest otherwise your anonymized names will change from one run to the other. We will see the difference later when we join to the well test table.

We could add the new ID column next to the original:

```{r}
# using the `apply` function to anonymize the `Wellname` column
well_data$well_id <- apply(well_data["Wellname"], 1, anonymize)  # 1: rows; 2: columns

# select only columns `Wellname` and `well_id`
well_data[c("Wellname", "well_id")]
```



```{r}
# using the function sapply
well_data$well_id <- as.vector(sapply(well_data["Wellname"], anonymize))
well_data
```

```{r}
# using the function apply
well_data$well_id <- apply(well_data["Wellname"], 1, anonymize)
well_data
```

```{r}
ncol(well_data)    # numbber of columns
names(well_data)   # names of the columns
```


```{r}
# move the well_id to the 1st column
well_data <- well_data[, c(17, 1:16)]
well_data
```




